---
title: "Lab 06-07"
author: "Debra Perrone"
date: "2/10/2021"
output: pdf_document
fontsize: 10pt
classoption: landscape
---


The data for this lab are available from the [United States Geological Survey](https://water.usgs.gov/watuse/data/index.html) for years 1950-2015. 
 
 

Table 1: Sets of data, for which each of the sets have ALL column tags in common. Data from USGS (1950-2015). 

year        | public supply           | irrigation              | rural                                             | industrial                                          | thermoelectric        | state
----        |----                     | ----                    | ----                                              | ----                                                | ----                  | ----
1950-1955   | ps_wgw_fr + ps_wsw_fr   | ir_wgw_fr + ir_wsw_fr   | **NA**                                            | **inpt_wgw_fr + inpt_wsw_fr**                       | **NA**                | **area**
1960-80     | ps_wgw_fr + ps_wsw_fr   | ir_wgw_fr + ir_wsw_fr   | do_wgwfr + do_wswfr + ls_wgwfr + ls_wswfr         | **oi_wsw_fr + oi_wgw_fr**                           | pt_wgw_fr + pt_wsw_fr | **area**
**1985-90** | ps_wgwfr + ps_wswfr     | ir_wgwfr + ir_wswfr     | **do_ssgwf + do_ssswf + ls_gwtot + ls_swtot**     | in_wgwfr + in_wswfr	+ **mi_gwtot + mi_swtot** | pt_wgwfr + pt_wswfr   | **scode**
1995        | ps_wgw_fr + ps_wsw_fr   | ir_wgw_fr + ir_wsw_fr   | do_wgw_fr + do_wsw_fr + ls_wgw_fr + ls_wsw_fr     | in_wgw_fr + in_wsw_fr + mi_wgw_fr + mi_wsw_fr       | pt_wgw_fr + pt_wsw_fr | **state_code**
2000        | ps_wgw_fr + ps_wsw_fr   | **it_wgwfr + it_wswfr** | do_wgw_fr + do_wsw_fr + ls_wgw_fr + ls_wsw_fr     | in_wgw_fr + in_wsw_fr +	mi_wgw_fr + mi_wsw_fr     | pt_wgw_fr + pt_wsw_fr | statefips
2005        | ps_wgw_fr + ps_wsw_fr   | ir_wgw_fr + ir_wsw_fr   | do_wgw_fr + do_wsw_fr + ls_wgw_fr + ls_wsw_fr     | in_wgw_fr + in_wsw_fr +	mi_wgw_fr + mi_wsw_fr     | pt_wgw_fr + pt_wsw_fr | statefips
2010-15     | ps_wgw_fr + ps_wsw_fr   | ir_wgw_fr + ir_wsw_fr   | do_wgw_fr + do_wswfr + **li_wgw_fr + li_wsw_fr**  | in_wgw_fr + in_wsw_fr +	mi_wgw_fr + mi_wsw_fr	  | pt_wgw_fr + pt_wsw_fr | statefips


### Step 1: Load libraries
```{r, warning = FALSE, message = FALSE, echo = FALSE}
library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(sf)
library(USAboundaries)
```

### Step 2: Read data
```{r, warning = FALSE, message = FALSE, error = TRUE, tidy = TRUE}

# Read data for 1950
wu_1950 <- lapply(excel_sheets(here("docs/data/data/us1950.xlsx")), 
              function(x) read_excel(here("docs/data/data/us1950.xlsx"), 
              skip = 3, 
              sheet = x)) %>%
        reduce(left_join, by = "Area") %>%
        clean_names() %>%
        mutate(across(c(1:9), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 1955
wu_1955 <- lapply(excel_sheets(here("docs/data/data/us1955.xlsx")), 
              function(x) read_excel(here("docs/data/data/us1955.xlsx"), 
              skip = 3, 
              sheet = x)) %>%
        reduce(left_join, by = "Area") %>%
        clean_names() %>%
        mutate(across(c(1:11), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 1960
wu_1960 <- lapply(excel_sheets(here("docs/data/data/us1960.xlsx")), 
              function(x) read_excel(here("docs/data/data/us1960.xlsx"), 
              skip = 3, 
              sheet = x)) %>%
        reduce(left_join, by = "Area") %>%
        clean_names() %>%
        mutate(across(c(1:35), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 1965
wu_1965 <- lapply(excel_sheets(here("docs/data/data/us1965.xlsx")), 
              function(x) read_excel(here("docs/data/data/us1965.xlsx"), 
              skip = 3, 
              sheet = x)) %>%
        reduce(left_join, by = "Area") %>%
        clean_names() %>%
        mutate(across(c(1:33), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 1970
wu_1970 <- lapply(excel_sheets(here("docs/data/data/us1970.xlsx")), 
              function(x) read_excel(here("docs/data/data/us1970.xlsx"), 
              skip = 3, 
              sheet = x)) %>%
        reduce(left_join, by = "Area") %>%
        clean_names() %>%
        mutate(across(c(1:34), as.numeric)) %>%
        replace(is.na(.), 0)


# Read data for 1975
wu_1975 <- lapply(excel_sheets(here("docs/data/data/us1975.xlsx")), 
              function(x) read_excel(here("docs/data/data/us1975.xlsx"), 
              skip = 3, 
              sheet = x)) %>%
        reduce(left_join, by = "Area") %>%
        clean_names() %>%
        mutate(across(c(1:34), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 1980
wu_1980 <- lapply(excel_sheets(here("docs/data/data/us1980.xlsx")), 
              function(x) read_excel(here("docs/data/data/us1980.xlsx"), 
              skip = 3, 
              sheet = x)) %>%
        reduce(left_join, by = "Area") %>%
        clean_names() %>%
        mutate(across(c(1:34), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 1985
wu_1985 <- read_delim(here("docs/data/data/us1985.txt"), delim = "\t") %>%
        clean_names() %>%
        mutate(across(c(2, 4:163), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 1990
wu_1990 <- read_excel(here("docs/data/data/us1990.xls"), sheet = 1) %>%
        clean_names() %>%
        filter(state != 3226) %>%
        mutate(across(c(2, 4:163), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 1995
wu_1995 <- read_excel(here("docs/data/data/us1995.xls"), sheet = 1) %>%
        clean_names() %>%
        mutate(across(c(1:4, 6:252), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 2000
wu_2000 <- read_excel(here("docs/data/data/us2000.xls"), sheet = 1) %>%
        clean_names() %>%
        mutate(across(c(2:70), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 2005
wu_2005 <- read_excel(here("docs/data/data/us2005.xls"), sheet = 1) %>%
        clean_names() %>%
        mutate(across(c(2:4, 6:108), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 2010
wu_2010 <- read_excel(here("docs/data/data/us2010.xlsx"), sheet = 1) %>%
        clean_names() %>%
        mutate(across(c(2, 4:117), as.numeric)) %>%
        replace(is.na(.), 0)

# Read data for 2015
wu_2015 <- read_excel(here("docs/data/data/us2015.xlsx"), sheet = 1, skip = 1) %>%
        clean_names() %>%
        mutate(across(c(2, 4:141), as.numeric)) %>%
        replace(is.na(.), 0)
```



```{r}
merge(wu_2010, wu_2015, by = 1:117)
'```













